---
title: "Nolan_Week_Two"
author: "Doug Nolan"
date: "12-08-2025"
output:
  pdf_document: default
  html_document: default
---
Assignment Details

1.Bar graph showing product and the sum of sales for each.

2.Pie chart showing sum of sales by region.

3.Box plots for average sales by region and average quantities by region.

4.Histogram of the count of orders by product.

5.A radial plot showing total sales (for all products) by month.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(knitr)  
library(tidyverse)
library(Ryacas)
library(gridExtra)
library(ggpattern)
library(ggplot2)
library(scales)
library(ggradar)

```

Import data from [Github](https://github.com/bellevue-university/dsc520/blob/master/assignments/version_two/10_11_week/week_02_and_03/sales_data_sample.csv)


```{r import data}
url <- "https://raw.githubusercontent.com/bellevue-university/dsc520/master/assignments/version_two/10_11_week/week_02_and_03/sales_data_sample.csv"

data <- read.csv(url)
## printing out data to validate it matches what is in github
head(data, n = 10)
dim(data) # rows and columns
summary(data) #basic stats
# data prep. get months organized (MDY format)
data$OrderDate <- mdy(data$OrderDate) # convert to date
data$Month <- month(data$OrderDate) # numeric month
data$Month_Label <- month(data$OrderDate, label = TRUE) # abbrev labels
```

Bar graph showing product and the sum of sales for each.
```{r bar-graph, echo=TRUE}
# Summarize sales by product
product_sales <- data %>%
  group_by(Product) %>%
  summarise(TotalSales = sum(TotalSales, na.rm = TRUE), .groups = 'drop')

# Bar graph
ggplot(product_sales, aes(x = Product, y = TotalSales, fill = Product)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = dollar_format()) +
  labs(title = "Total Sales by Product", x = "Product", y = "Total Sales ($)") +
  theme_minimal() +
  theme(legend.position = "none")

```

Pie chart showing sum of sales by region.
```{r pie-chart}
sales_by_region <- data %>%
  group_by(Region) %>%
  summarise(total_sales = sum(TotalSales, na.rm = TRUE)) %>%
  mutate(
    pct = total_sales / sum(total_sales),
    label = paste0(Region, "\n$", scales::comma(total_sales))
  )
# pie chart
ggplot(sales_by_region, aes(x = "", y = total_sales, fill = Region)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  geom_label(aes(label = label),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE,
             label.size = 0) +
  labs(title = "Sales Distribution by Region") +
  theme_void()


```


Box Plots - Total Sales and quantity by Region
```{r box plot}
#box plot for total sales by region
p1 <- ggplot(data, aes(x = Region, y = TotalSales, fill = Region)) +
  geom_boxplot() +
  scale_y_continuous(labels = dollar_format()) +
  labs(title = "Distribution of Total Sales by Region", x = "Region", y = "Total Sales ($)") +
  theme_minimal() +
  theme(legend.position = "none")

# box plot for quantity by region
p2 <- ggplot(data, aes(x = Region, y = Quantity, fill = Region)) +
  geom_boxplot() +
  labs(title = "Distribution of Total Sales by Region", x = "Region", y = "Quantity") +
  theme_minimal() +
  theme(legend.position = "none")

grid.arrange(p1, p2, ncol = 2)
```

Histogram of the count of orders by product
```{r histogram}
# count orders by product
product_count <- data %>%
  group_by(Product) %>%
  summarise(Order_Count = n(), .groups = 'drop')

# bar plot / histogram
ggplot(product_count, aes(x = Product, y = Order_Count, fill = Product)) +
  geom_bar(stat = "identity") +
  labs(title = "Number of Orders by Product", x = "Prodcut", y = "Order Count") + 
  theme_minimal() +
  theme(legend.position = "none")
```






A radial plot showing total sales (for all products) by month
```{r radial plot2}
# Prepare monthly data

monthly_sales <- data %>%
mutate(Month = month(OrderDate, label = TRUE)) %>%
group_by(Month) %>%
summarise(total_sales = sum(TotalSales, na.rm = TRUE))

# Normalize for radar chart

radar_data <- monthly_sales %>%
mutate(total_sales_scaled = total_sales / max(total_sales)) %>%
select(Month, total_sales_scaled)

# Convert to wide format for ggradar

radar_wide <- radar_data %>%
pivot_wider(names_from = Month, values_from = total_sales_scaled)

# ggradar requires a data frame with rownames

radar_df <- as.data.frame(radar_wide)
row.names(radar_df) <- "Monthly Sales"

# Create radar chart

ggradar(radar_df,
values.radar = c("0","0.5","1"),
grid.min = 0,
grid.mid = 0.5,
grid.max = 1,
legend.position = "none") +
ggtitle("Radial Plot of Total Monthly Sales")


```


